---
title: "R Notebook"
output: html_notebook
---


# Question 1 (10 marks)
The data set olympicbutterfly.csv contains the winning time in seconds for the
men’s and women’s 100m butterfly race for the Olympics from 1964 through 2016.
Let yi and xi denote the winning time and year for the ith Olympics. In addition,
let wi denote an indicator variable that is 1 for the women’s race and 0 for the men’s
race.
Consider the regression model
Yi ∼ Normal(μi, σ2), μi = β0 + β1(xi − 1964) + β2wi.
(a) (2 marks) Interpret the meaning in terms of the winning time for i) β0 + β2;
ii) β0 + 10β1 + β2, respectively.


(b) (4 marks) Suppose the regression parameters β0, β1, β2 and the precision pa-
rameter 1/σ2 are assigned weakly informative priors. Using JAGS, sample 5000
draws from the joint posterior distribution of all parameters.
## Import packages
```{r}
library(rjags)
library(tidymodels)
library(coda)
```

## Import dataset
```{r}
data_1 <- olympicbutterfly
data_1
```
## Simple Data Exploration
```{r}
data_1 %>%  skimr::skim()
```
## Define the values for the simulation
```{r}
#Values for simulation
N <- length(data_1$Year)
N
```


## Data Preparation
```{r}
x <- data_1$Year - 1964
w <- ifelse(data_1$Gender == "Women", 1, 0)
y <- data_1$Time
```


## Combine data for JAGS input
```{r}
jags_data <- list("N" =N ,"x"=x, "w" =w, "y"= y)
jags_data
```


```{r}
#Initial values
inits <- list( d = 0, tau = 1, delta = c (0,0,0,0,0, 0,0), mu = c (0,0,0,0,0, 0,0) )
```

## Embed the BUGS model 
# Define the JAGS model

```{r}
cat("
model {
  for (i in 1:N) {
    y[i] ~ dnorm(mu[i], tau)
    mu[i] <- beta[1] + beta[2] * x[i] + beta[3] * w[i]
  }

  beta[1] ~ dnorm(0, 0.001)  # Weakly informative priors
  beta[2] ~ dnorm(0, 0.001)
  beta[3] ~ dnorm(0, 0.001)
  tau ~ dgamma(0.001, 0.001)  # Weakly informative prior for precision 

 
}
", file="model8.txt")
```


## Specify the JAGS model
```{r}
## Set up the JAGS model
jags.m <- jags.model( file = "model8.txt", data= jags_data, inits=inits, n.chains=1, n.adapt=5000 )
```

## Specify the parameters of interest.
```{r}
## specify parameters to be monitored
params <- c("d", "OR", "tau")
```

## Run JAGS model and save the samples
```{r}
## run JAGS and save posterior samples
samps <- coda.samples( jags.m, params, n.iter=10000 )
```

## Use a burn-in of 5,000 iterations
```{r}
## summarize posterior samples
summary(samps)
```

## 
```{r}
summary(window(samps, start=5001))  # Burn in of 5000. Start at 5001.
```

## Plot the posterior distribution
```{r}
plot(samps)
```


(c) (4 marks) Using sample draws from part (b), construct 90% interval estimates
for each of the regression coefficients. Based on your work, describe how the
mean winning time in the butterfly race has changed over time. In addition,
describe how the men’s times differ from the women’s times


## Question 2(c)
```{r}
gibbs_sampler <- function(iterations, a, b) {
  Y <- rep(0, iterations)
  lambda <- rep(0, iterations)
  
  # Initial values
  Y[1] <- rpois(1, lambda[1])
  lambda[1] <- rgamma(1, a, b)
  
  for (i in 2:iterations) {
    # Sample Y from the conditional distribution Y|lambda
    Y[i] <- rpois(1, lambda[i - 1])
    
    # Sample lambda from the conditional distribution lambda|Y
    lambda[i] <- rgamma(1, a + Y[i], b + 1)
  }
  
  return(list(Y = Y, lambda = lambda))
}
```

## Question 2(d)
(d) (2 marks) Use the R function in part (c) to run 1000 iterations of Gibbs
sampling for the case where a = 3 and b = 3. Report your simulated draw of
Y and corresponding relative frequencies.
```{r}
# Run Gibbs sampling for 1000 iterations for the case where a = 3 and b = 3
outcome <- gibbs_sampler(iterations = 1000, a = 3, b = 3)

# Simulated draw of Y
Y_draw <- outcome$Y

# Relative frequencies of Y
relative_frequencies <- table(Y_draw) / sum(table(Y_draw))
relative_frequencies

```












